
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# Install from GitHub

*Written by Haoluan Chen and last updated on 7 October 2021.*

## Introduction

In this lesson, you will learn how to install package from GitHub.

Prerequisite skills include:

- Having RStudio set-up and working.
- Comfort running R code in the console.
- Comfort installing package from CRAN.

## What is GitHub?

GitHub is a popular repository for open-source projects. Many developers can share and collaborate their work with the world. Sometimes, you may need to use the package from GitHub for a specific task.

Note: Unlike the packages from CRAN, there is no review process for GitHub packages, so you may encounter bugs or errors when using them.  

## How to install packages from GitHub?

To install the package from GitHub, you will need the function `install_github()` in the `devtools` package. `devtools` contains tools for package developers to build and test their packages. 

So, you need to first install `devtools` package and load the package in order to use the `install_github` function. However, a faster way to do that is to run the following code in the r chunk or the console: `devtools::install_github("developer name/package name")`. The developer name and package name can be found on the top left side of the GitHub repository

This code means that you want to call the `devtools` package and use the function `install_github()`. This way, you do not have to install and load the `devtools` package.

## Example

https://github.com/sharlagelfand/dmc

The goal of `dmc` package is to allow you to find the closest DMC embroidery floss colour(s) for a given colour, as well as access colour (hex, RGB) information about DMC colours.

You can run the following code on your console to install this package: 

```{r eval = FALSE}
devtools::install_github("sharlagelfand/dmc")
```

After running this code, you will receive some message on the console and it will start installing the package (this may take a few minutes depending on your system). Then it will tell you if you successfully installed the package.

Note: for some packages, it may ask you to input something in the console to complete the installation. For `dmc` package, you will see "These packages have more recent versions available. Which would you like to update?" and "Enter one or more numbers, or an empty line to skip updates:" 

<iframe width="560" height="315" src="https://www.youtube.com/embed/hDQKBBWu98k" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

## Exercises

### Exercise 1

Please install the following package on your RStudio and answer the question

https://github.com/RohanAlexander/AustralianPoliticians


<!-- ```{r githubex1, echo = FALSE} -->
<!-- question("Which code can install above package?", -->
<!--           answer("devtools::install_github('RohanAlexander/AustralianPoliticians')", correct = TRUE), -->
<!--           answer("devtools::install('RohanAlexander/AustralianPoliticians')"), -->
<!--           answer("install_github('RohanAlexander/AustralianPoliticians')"), -->
<!--           answer("devtools::install_github('AustralianPoliticians')"), -->
<!--           answer("devtools::install_github('RohanAlexander')"), -->
<!--           allow_retry = TRUE) -->

<!-- ``` -->

<iframe width="560" height="315" src="https://www.youtube.com/embed/WlnawRIELzY" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

## Common mistakes and errors

- Pay attention to R console, it may ask you to input something
- Make sure the repository is a R package
- Make sure the spelling of the developer name and package name are correct
- Make sure spelling of the function is correctly called devtools::install_github("developer name/package name")

## Next steps

For next step, you can install the package from GitHub and start using them for your projects!

In the future, you may want to build your own R package!

R Packages is a book that gives a comprehensive treatment of all common parts of package development and uses `devtools` throughout. The first edition is available at 'http://r-pkgs.had.co.nz', but note that it has grown somewhat out of sync with the current version of `devtools`.

## Exercises

### Question 1

### Question 2

### Question 3

### Question 4

### Question 5

### Question 6

### Question 7

### Question 8

### Question 9

### Question 10
