
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(flair)
library(glue)
library(stringr)
library(palmerpenguins)
```

# paste, paste0, glue and stringr

*Written by Marija Pejcinovska and last updated on 7 October 2021.*


## Introduction  

In this lesson we'll cover a couple of ways in which you can paste strings to your data. We'll use the `paste()` and `paste0()` functions, both of which are part of base R,  as well as the `glue()` function in the `glue` package. As part of this tutorial, you will also get a first taste of the `stringr` package, which is a wonderful tool for handling strings.  

Prerequisite skills:

- Exposure to most of the material in the yellow-level tutorials would be helpful.



## Paste strings and data with `paste()` and `paste0()`

`paste()` and `paste0()` are handy R functions used for concatenating strings and/or data. One of the main difference between the two is the default setting in their arguments, specifically the argument used to denote the way in which results should be separated. Both `paste()` and `paste0()` convert all passed objects to character vectors.   

Both functions have the following two arguments: `sep = ` and `collapse = `.

Let's see `paste()` in action first.   

You can pass individual objects to `paste()`
```{r}

paste(1, "a", "b")

```
 or pass vectors.
```{r}
paste(1:3, c("a", "b", "c"))
```
Notice that when the arguments passed are vectors, the concatenation happens term-by-term. Vector arguments are recycled as necessary.  
For instance, note the difference between
```{r}
paste(1:2, "a", "b", "c")
```
and 
```{r}
paste(1:2, c("a", "b", "c"))
```
You can see that in the latter example the concatenation is term-by-term and the sequence `1:2` is getting recycled (i.e. restarted) to match the length of `c("a", "b", "c")`.  

The argument `sep = ` in the function controls which character string is used to separate the terms. By default, the `paste()` function sets `sep = " "`, meaning concatenated terms are separated by empty space. We can set the separator to be any string character we'd like. 

```{r}
## Leaving no space
paste(1:2, c("a", "b", "c"), sep = "")

## Concatenating with a dash
paste(1:2, c("a", "b", "c"), sep = "-")

## Concatenating with a random letter
paste(1:2, c("a", "b", "c"), sep = "Y")

##..and so on
```

If a value is specified for the argument `collapse = `, the elements in the result are then turned into a single string, with the components being separated by the character string provided in `collapse`. For instance, we can turn 
```{r}
paste(1:2, c("a", "b", "c"), sep = "-")
```
into a single string as
```{r}
paste(1:2, c("a", "b", "c"), sep = "-", collapse = ", ")
```

where we've used `,  ` ` ` (comma with a blank space) as a separator.

The most preferred separator tends to be the "no space" one. This is the default setting of `paste0`.

Run the following code to see the difference between `paste` and `paste0`

```{r ex1_paste_, exercise=TRUE, exercise.lines=5}

paste("I", "Love", "R","!")

paste0("I", "Love", "R", "!")


```

Of course, you might be interested in a more sophisticated collection of objects to paste. For instance, suppose you needed to work with a combination of character strings and existing R objects.

```{r student-example-paste}
student <- c("Rohan", "Monica")
badges <- c(2,4)
term <- c("Fall", "Spring")

paste(student, "collected", badges, "DoSS toolkit badges in the", term )

```


## "Gluing" your data with the `glue` package

The `glue` package is designed to make it easier to "stitch" or interpolate (or "glue") your data into strings. Its main function is quite similar in flavor to `paste()` (and `paste0()`), but a bit easier to use (this is especially true when compared to `sprintf()`; a function that we have not discussed here, but can similarly be used to concatenate strings and data).  

You can download `glue` the usual way you would install all your other packages, that is, by calling `install.packages("glue")`.

### The functions of `glue`
The glue package has three primary functions, `glue()`, `glue_data()` and `glue_collapse()`.   

The **`glue()` function** works a bit like the `paste()` function. In the case of `glue()`, however, we use `{}` to wrap the R code we wish to reference inside the string. This makes it a bit more manageable compared to all the quotation marks, commas, and separators we need to keep track of when using `paste()`.

Let's refer back to our example from earlier and check out the syntax for `glue()`. Note that now everything is placed inside a single set of quotation marks and R objects are referenced within the string by wrapping them up in curly brackets. 

```{r student-example-glue}
student <- c("Rohan", "Monica")
badges <- c(2,4)
term <- c("Fall", "Spring")

glue("{student} collected {badges} DoSS toolkit badges in {term}")

```

Here we've called the R objects `student`, `badges`, and `term` by wrapping them up in `{}`. 

If you wish to use something other than `{}`, you can specify different opening and closing delimiters by using the `.open = ` and `.close = ` arguments in the `glue` function. For instance, let's surround the code we wish to evaluate by `<` at the beginning and `]` at the end.

```{r student-example-glue2}
student <- c("Owen", "Monica")
badges <- c(2,4)
term <- c("Fall", "Spring")

glue("<student] collected <badges] DoSS toolkit badges in <term]", .open = "<", .close = "]")

```

The function **`glue_data()`** works much like `glue()` but it is designed to be used in piped chains (recall the pipe operator, ` %>% `). What's important to note here is that inside the curly braces we pass the column names of the columns in our data we wish to glue in some way (and not the name of the variables or R objects as we did with `glue()`). 

As an example run the following code. See what happens if you change `glue_data()` below with `glue()`.

```{r penguin-example-glue-data, exercise = TRUE, exercise.lines=12}

library(palmerpenguins)
# Check the data out
penguins %>% 
  head()

# "Glue" some data elements together
penguins %>% 
  slice(1:4) %>% 
  glue_data("This {species} penguin living on {island} \\
            island has flipper length {flipper_length_mm} mm") 

# note: the "\\" symbol in the string code above is simply to force the string to appear as one line. 

```


Finally, the **`glue_collapse()` function** concatenates multiple values into one. This function has a particularly clever and useful argument called `last` which allows you to change the separator for the last value (a feature not available in `paste()`).

```{r}
glue_collapse({letters[1:3]}, sep = ", ", last = ", and ")

```


## First taste of the `stringr` package

As a budding data scientist, you may have discovered that data analysis tasks usually involve spending an outsize portion of your time cleaning and processing data (it's a very special tax you have to pay before you get to the "fun" data scienc-y bits and pieces!!).   

On occasion your data may contain lots of text or strings. The `stringr` package, which is part of the core *tidyverse*, is a wonderful collection of functions that make string manipulation easier.   

This section is intended to get you started with the `stringr` package and involves just a brief introduction to the `str_detect()` and `str_replace()` functions in `stringr`.   

All functions in this package begin with the `str_` prefix and have easy to remember, intuitive names. 

For instance, `str_detect()` allows you to check whether a vector of characters contains a particular pattern, in other words, it allows you to *detect* a pattern in a string. The function returns a logical vector of the same length as the input, where `TRUE` indicates the pattern has been matched in that index. 

Let's check this out.
```{r str-detect-1}

my_vec <- c("r", "R", "I love R", "why?", "what is this?")

my_vec %>% 
  str_detect("r")

my_vec %>% 
  str_detect("wh")
```

Note that the pattern detection is case sensitive. 

`str_replace()`, on the other hand, allows you to replace a matched pattern with an entirely new string (or an empty string, if what you  wish to do is remove the pattern; see also `str_remove()`). 
```{r str-remove-1}
my_vec 

my_vec %>% 
  str_replace("\\?", " code")

```

The first argument is the pattern we are trying to find a match for (note that since ? is a special character we need to escape it by placing `\\` before `?`). The second argument here is the string we wish to replace our matched pattern with (in our case that's the string "code"). 

Let's see the `str_replace()` function applied to our penguins data. We'll use the function to slightly modify the name of an island

```{r}

penguins %>% 
  slice(1:4) %>% 
  mutate(new_island = str_replace(island, "Tor", "Mor"))
  
```


This is just a small slice of what the `stringr` package has to offer. Be on the lookout for future tutorials with more on strings and regular expressions. 


## Exercises


Consider the first few rows of the penguin data we saw in an example earlier. For this exercise we'll call the sliced data `small-penguins`. To refresh your memory of what the data frame looks like it's been reproduced here (use the arrows to navigate left and right through the columns). 

```{r small_penguins, echo=TRUE}
small_penguins <- penguins %>% 
  slice(1:10)
small_penguins
```



```{r small_penguins0, echo=TRUE}
small_penguins <- penguins %>% 
  slice(1:7)
small_penguins
```

### Exercise 1

Change this function from the `glue` package to produce the following string "This penguin is of the species XXX from island XXX and has body mass of XXX g" describing the 6th penguin in `small_penguins`, where the three XXX should be species, island name, and body mass respectively.


```{r ex1_glue_data, exercise=TRUE, exercise.lines=8, exercise.eval=FALSE, exercise.setup = "small_penguins"}

small_penguins %>% 
  slice(6) %>% 
  glue_data("This penguin is of the species {island} from island {body_mass_g} and has body mass of {species} g" )
  

```

```{r ex1_glue_data-solution}

small_penguins %>% 
  slice(6) %>% 
  glue_data("This penguin is of the species {species} from island {island} and has body mass of {body_mass_g} g" )
  
```



### Exercise 2

Using `glue()` and `glue_collapse()` manipulate the vectors `quant` and `food` below to get the following single string "1 sushi roll, 2 tacos, and 3 cakes"

```{r ex2_glue_collapse, exercise=TRUE, exercise.lines=8, exercise.eval=FALSE}

quant <- 1:3
food <- c("sushi roll", "tacos", "cakes")


```

```{r ex2_glue_collapse-solution}

quant <- 1:3
food <- c("sushi roll", "tacos", "cakes")

glue_collapse((glue("{quant} {food}")), sep = ", ", last = ", and ")

# Or alternatively with the piping syntax
glue("{quant} {food}") %>% 
  glue_collapse(.,sep = ", ", last = ", and")
  
```


### Exercise 3 

Back to our `small_penguins` data from Exercise 1.

Use `str_detect()` and `filter()` to subset the `small_penguins` data to include only female penguins. 

```{r ex3_str-detect, exercise=TRUE, exercise.lines=8, exercise.eval=FALSE, exercise.setup = "small_penguins"}
small_penguins %>% 
  filter(str_detect(sex, "male"))
```

```{r ex3_str-detect-solution}

small_penguins %>% 
  filter(str_detect(sex, "fe"))
  
```





### Exercise 4

Use  `mutate()` to create a new variable `island_short` where with the help of `str_replace` you remove the pattern "ersen" from the island name.

```{r ex4_str-remove, exercise=TRUE, exercise.lines=8, exercise.eval=FALSE, exercise.setup = "small_penguins"}

 small_penguins %>% 
  mutate(island_short = str_replace(island, pattern = "island", replacement = ""))

```

```{r ex4_str-remove-solution}

 small_penguins %>% 
  mutate(island_short = str_replace(island, pattern = "ersen", replacement = ""))

```




## Next Steps


This tutorial only briefly introduced you to the capabilities of the `stringr` package. Future tutorials will revisit this package and dive deeper into some of its functions. In the meantime, you can learn more about this package in Hadley Wickham's book *"R for Data Science"*. The link here takes you to the chapter on strings which also introduces you to regular expressions (regular expressions are a bit tedious, but you might learn to like them if you end up working with text data a lot). https://r4ds.had.co.nz/strings.html 



## Exercises

### Question 1

### Question 2

### Question 3

### Question 4

### Question 5

### Question 6

### Question 7

### Question 8

### Question 9

### Question 10
