
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(flair)
library(glue)
library(stringr)
library(palmerpenguins)

broadway <- read_csv(
  "https://raw.githubusercontent.com/tacookson/data/master/broadway-grosses/grosses.csv",
  guess_max = 10000)

broadway <- broadway %>%
  select(week_ending, show, theatre, weekly_gross, avg_ticket_price,
         top_ticket_price, performances, previews)

```


# Conditional mutating and summarising

*Written by Mariam Walaa.*

## Introduction

In this lesson, you will learn how to:

- Use `across()` with `summarise()` 
- Use `mutate_if()` 
- Use `if_else()` and `na_if()` 

Prerequisite skills include:

- Familiarity with `summarize()` and `mutate()`
- Familiarity with conditional statements `if_else()`

Highlights:

- Use `across()` to summarize across a defined selection of columns
- Mutate column types based on conditions using `mutate_if()`
- Mutate columns based on conditions using `if_else()` and `na_if()` within `mutate()`

## Overview

This section will demonstrate how to use the `summarise()` function with `across()` to 
summarize variables and groups within a variable in a data set. We will be looking at a
data set of Broadway shows with variables about the performances, attendance, and revenue
for theaters that are part of The Broadway League. You can learn more about the data set
provided by Alex Cookson in [the dataset repository](https://github.com/tacookson/data) as
well as this corresponding [blog
post](https://www.alexcookson.com/post/most-successful-broadway-show-of-all-time/).

### Video

![](https://youtu.be/3mmbPjgzzeY)

## Questions

Lets start with loading the tidyverse and looking at the data.

```{r load-tidyverse}
library(tidyverse)
```

```{r across-2}
broadway
```

### Question 1

What is the minimum and maximum number of performances _and_ previews per week? We can use
`across()` to select specific columns to summarize them with multiple summary functions.

```{r across-3, include = FALSE}
broadway %>%
  group_by(week_ending) %>%
  summarise(across(.cols = c("performances", "previews"),
                   .fns = list(min = min, max = max)),
            .groups = 'drop')
```
```{r across-4, echo = FALSE}
decorate("across-3") %>%
  flair("across", color = "blue") %>%
  flair(".cols", color = "deeppink") %>%
  flair(".fns", color = "orange") %>%
  knit_print.with_flair()
```

Here is what the above chunk of code does:

* Groups the data by week using `group_by()` 
* Selects columns to summarize by passing a vector to `.cols` in `across()` (Highlighted in pink) 
* Defines summary functions by passing a list to `.fns` in `across()` (Highlighted in orange)

You can also learn more about `across()` by running `?across` in your console.

### Question 2

How do we provide a numeric summary for every show happening in a particular week?

```{r across-5}
broadway %>%
  group_by(week_ending, show) %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE),
            .groups = 'drop')
```

This chunk of code summarizes every show happening in some particular week by every
numeric variable that is available in the data set. This helps us compute things like
the average ticket price and the number of performances each show had in a particular
week.

### Question 3

We can use mutate() and if_else() to change values within a column. For example, here
is what we would write if we wanted to change all rows with theatre Studio 54 to be 
called Studio 54 Theatre.

```{r across-6}
broadway %>% 
    mutate(theatre = if_else(theatre == "Studio 54", "Studio 54 Theatre", theatre))
```

We can confirm this by filtering for these theatre names.

```{r across-7}
broadway %>%
  mutate(theatre = if_else(theatre == "Studio 54", "Studio 54 Theatre", theatre)) %>%
  filter(theatre %in% c("Studio 54", "Studio 54 Theatre"))
```

Notice that we did not save the initial modified data frame. We are also using a new
operator %in% which checks for the value Studio 54 in a vector of values including
Studio 54 and Studio 54 Theatre.

## Arguments

### `across()`

The `across()` function takes the following as arguments:

| Argument | Parameter | Details                                           |
| -------- | --------- | ------------------------------------------------- |
| .fns     |           | can pass a list of functions or a single function |
| .cols    | vector    | vector with column names to apply functions to    |

You can read more about the arguments in the `across()` function documentation
[here](https://dplyr.tidyverse.org/reference/across.html). Please note that the arguments 
for across() will also depend on your use case.

### `mutate_if()`

The `mutate_if()` function takes the following as arguments:

| Argument  | Details                                            |
| --------- | -------------------------------------------------- |
| selection | selection of variables by type, i.e., is.character |
| function  |  function to apply to the selection of variables   |

You can read more about the arguments in the `mutate_if()` function documentation
[here](https://dplyr.tidyverse.org/reference/mutate_all.html). 

### `na_if()`

The `na_if()` function takes the following as arguments:

| Argument  | Details                                            |
| --------- | -------------------------------------------------- |
| vector    | vector to look for values we want to change        |
| value     | value in the vector that we want to change to NA   |

You can read more about the arguments in the `na_if()` function documentation
[here](https://dplyr.tidyverse.org/reference/na_if.html).

You can read more about `if_else()` in the corresponding tutorial on conditional
statements.

## Exercises

### Exercise 1

Count the number of distinct shows and distinct theatres using summarise() and across().
As a tip, try to use the data type for shows and theatres columns.

```{r across-exercise-1, exercise = TRUE}

```

```{r across-exercise-1-sol, echo = FALSE}
across_ex1_sol <- broadway %>%
  summarise(across(where(is.character), n_distinct),
            .groups = 'drop')
```

### Exercise 2

Use the mutate_if() function to select all the variables of type double and convert them
to character.

```{r across-exercise-2, exercise = TRUE}

```

```{r across-exercise-2-sol, echo = FALSE}
across_ex1_sol <- broadway %>% mutate_if(is.double, as.character)
```

### Exercise 3

There are some shows with a value of [title of show] under the show column. Use the
na_if() conditional statement within mutate() to convert this to NA.

```{r across-exercise-3, exercise = TRUE}

```

```{r across-exercise-3-sol, echo = FALSE}
across_ex1_sol <- broadway %>%
    mutate(show = na_if(show, "[title of show]"))
```


## Next Steps

If you would like to learn more about the functions in this tutorial, you may find the
following resource helpful:

- [dplyr: Apply a function (or functions) across multiple
columns](https://dplyr.tidyverse.org/reference/across.html)
- [Mutate multiple columns](https://dplyr.tidyverse.org/reference/mutate_all.html)
- [Convert values to NA](https://dplyr.tidyverse.org/reference/na_if.html)










