
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(flair)
library(glue)
library(stringr)
library(palmerpenguins)
```

# Working with dates

*Written by Mariam Walaa and last updated on 7 October 2021.*

## Introduction

In this lesson, you will learn how to:

- Work with dates using the `lubridate` library

Prerequisite skills include:

- Installing and loading packages
- Using `mutate()` and `summarise()`

Highlights:

- Loading the `lubridate` library separately from `tidyverse`
- Using the `today()` and `now()` functions
- Learning how to extract information from a date-time column


## Overview

Let's start by loading the `tidyverse`.

```{r dates-1, eval = TRUE}
library(tidyverse)
```

Notice that `lubridate` is not listed as one of the libraries loaded into your R session
when you load `tidyverse` (which you can view in your output on your own RStudio), and
this is because it is not part of the core `tidyverse` so it will need to be loaded
separately.

```{r dates-2}
library(lubridate)
```

Two of the functions the `lubridate` library will give us is `today()` and `now()`, which
we can immediately start to use, without data or parameters.

```{r dates-3}
today()
```

```{r dates-4}
now()
```

You may want to use these functions today() and now() for various reasons, such as 
including the date and time in a file name. You can do that as follows.

```{r}
paste0(today(), "_datafile.csv")
```

### Example

We will look at Caribou data from Alex Cooksons dataset repository for this tutorial.

```{r dates-5, include = FALSE}
caribou <-
  read_csv("https://raw.githubusercontent.com/tacookson/data/master/caribou-location-tracking/locations.csv")
```

```{r dates-6}
glimpse(caribou)
```

We can see that the timestamp column has an associated data type 'dttm.' This stands for
datetime. Given that this is already the correct data type, we can extract a lot of
information from it using `lubridate` functions.

```{r dates-7}
# extracting year
caribou %>% mutate(year = year(timestamp))

# extracting week day
caribou %>% mutate(week_day = wday(timestamp))

# extracting whether it's a leap year!
caribou %>% mutate(leap_year = leap_year(timestamp))
```

You can see there are many additional variables we can extract with the `lubridate` 
package. You can check out the [lubridate
cheatsheet](https://raw.githubusercontent.com/rstudio/cheatsheets/master/lubridate.pdf) to
learn more.

## Exercises

### Exercise 1

Extract month, day, and year into separate columns at the same time using mutate.

```{r dates-exercise-1, exercise = TRUE}

```

```{r dates-exercise-1-sol, echo = FALSE}
dates_ex1_sol <- caribou %>% mutate(month = month(timestamp),
                              day = day(timestamp),
                              year = year(timestamp))
```


### Exercise 2

Find the earliest and latest dates in the timestamp column.

```{r dates-exercise-2, exercise = TRUE}

```

```{r dates-exercise-2-sol, echo = FALSE}
dates_ex2_sol <- caribou %>% summarise(min(timestamp), max(timestamp))
```


**Hint**: Remember the summary functions you learned in `summarize()` tutorial!

## Common Mistakes & Errors

- You might try to apply `lubridate` functions on to a column that looks like a timestamp
column, but it may still not be of the appropriate data type, which is either 'dt', 'tm',
or 'dttm'. Be sure to convert to the proper data type first!

## Next Steps

If you would like to learn more about handling dates and times in R, as well as the
`lubridate` package, please read the following:

- [R For Data Science:  16 Dates and times](https://r4ds.had.co.nz/dates-and-times.html)
- [Lubridate Package Documentation](https://lubridate.tidyverse.org/)




## Exercises

### Question 1

### Question 2

### Question 3

### Question 4

### Question 5

### Question 6

### Question 7

### Question 8

### Question 9

### Question 10
