
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(flair)
library(knitr)
```

# Overview of ggplot2

*Written by Shirley Deng and last updated on 7 October 2021.*

## Introduction to `ggplot()` through the scatterplot

Have you ever made a plot in base R? It's not very pretty. 

Consider the **scatterplot**. 

### What is a scatterplot?

In a scatterplot, each observation is graphed as a point on a grid with axes x and y. 

We determine the position of each point based on the values of two numerical variables: the horizontal x-axis, and the vertical y-axis. 

### Let's look at an example of a scatterplot

We can make one right now using the built-in `mtcars` dataset to take a look - below is a scatterplot made with base R.

``` {r, ggplot-base-ex}
plot(x = mtcars$wt, 
     y = mtcars$mpg,
     main = "Miles vs. Weight", 
     xlab = "Car Weight (1000 lbs)", 
     ylab = "Miles per US Gallon",
     pch = 16
)
```

Alternatively, we can make the same plot with the `ggplot2` package, using `ggplot()` and `geom_point()`.

``` {r, ggplot-alt-ex}
mtcars %>% ggplot(aes(x = wt, y = mpg)) +
  geom_point() +
  labs(title = "Miles vs. Weight", 
       x = "Car Weight (1000 lbs)", 
       y = "Miles per US Gallon") +
  theme_classic() +
  theme(plot.title = element_text(hjust=0.5))
```

But `ggplot2` also allows us to customize our plots, and even plot them together as one figure.

``` {r, ggplot-alt-ex-pretty}
library(ggpubr)

base_plot <- mtcars %>% 
  ggplot(aes(x = wt, y = mpg))

grey_plot <- base_plot +
  geom_point(color = "#7570b3") +
  labs(x = "Car Weight (1000 lbs)", y = "Miles per US Gallon", 
       caption = "Using theme_grey()") +
  theme_grey()

minimal_plot <- base_plot + 
  geom_point(color = "#66a61e") +
  labs(x = "Car Weight (1000 lbs)", y = NULL, 
       caption = "Using theme_minimal()") +
  theme_minimal()

theme_plots <- ggarrange(grey_plot, minimal_plot)
theme_plots_title <- text_grob("Miles vs. Weight", face = "bold")
annotate_figure(theme_plots, top = theme_plots_title)
```

Looks cleaner, right? Or at least more fun.

## Using `ggplot()`

### What is `ggplot()` for?

`ggplot2` is a package that allows us to make graphics in R. It's a part of the `tidyverse`, so when you load the `tidyverse` package, you've loaded `ggplot2` too.

```{r ggplot-setup, eval=FALSE}
library(tidyverse)
```

The `ggplot2::ggplot()` function is used to initialize plots in R. We use it to indicate the base of our plot - what **data** we're working with, and any plot **aesthetics** we'll be using.

From there, we can add on layers to specify what the actual plots will be. This includes information on the **type of graph** we want, what we want its **axes** to look like, what we want its **labels** to look like, etc. 

It may seem farfetched now, in the middle of a *panoramic*, but imagine going to a restaurant to eat. 

You'd get seated in the restaurant you're going to eat in before you start to order food. We can think of `ggplot()` as getting seated in the restaurant, and adding additional layers as ordering food. You can add as many layers as you want at any time, just like how you can also order as many menu items as you want at any time.


### What arguments does `ggplot()` take?

* `.data`: the dataframe we're working with as a base
* `.mapping`: the aesthetics we're working with as a base - more on this below.

### What are aesthetics?

That's a separate course in philosophy. 

But, for our purposes, this argument is used to specify **which variables** of the dataframe we want to use for our plot's axes, among other options.

For example, with our `mtcars` example from earlier, we wanted the `wt` variable on the x-axis and the `mpg` variable on the y-axis, so we used `mapping = aes(x = wt, y = mpg)`.

#### Exercise 1

What if we wanted to switch the axes, with `mpg` on the x-axis and `wt` on the x-axis? Try modifying the code below to generate the plot.

```{r ggplot-exercise1, exercise=TRUE}
mtcars %>% 
  ggplot(aes(x = wt, y = mpg)) +
  geom_point() +
  labs(title = "Weight vs. Miles", 
       x = "Miles per US Gallon",
       y = "Car Weight (1000 lbs)")
```

```{r ggplot-exercise1-hint}
aes(x = mpg, y = wt)
```

### What else do we need to use `ggplot()` ? **Layers**

Once we've set our `ggplot()` aesthetics, we need to specify what exactly we want to plot by using **layers**. This includes the **type of plot** - we use geom layers to do so.

#### Geoms

Some common geoms include:

* `geom_bar()` - bar plot
* `geom_boxplot()` - box and whiskers plot
* `geom_histogram()` - histogram
* `geom_point()` - scatterplot

In Exercise 1 and our previous examples, we used `geom_point()` to specify that we wanted to plot a scatterplot.

#### Example 1 - the `+`

Let's take a look at the following example:

```{r ggplot-example1, eval=FALSE}
mtcars %>% ggplot(aes(x = wt, y = mpg)) +
  geom_point() +
  labs(title = "Miles vs. Weight", 
       x = "Car Weight (1000 lbs)", 
       y = "Miles per US Gallon")
```

After we called `ggplot()`, we used a `+` to **connect** it with our `geom_point()`.

Similarly, we added another layer - **labels**. We used another `+` to connect the labels `labs()` with the rest of the figure.

#### Themes 

There are also layers for changing the **appearance** of our plots. For example, we can use the **theme** `theme_void()` to remove the axes and axes' labels:

```{r ggplot-example1-void}
mtcars %>% ggplot(aes(x = wt, y = mpg)) +
  geom_point() +
  labs(title = "Miles vs. Weight", 
       x = "Car Weight (1000 lbs)", 
       y = "Miles per US Gallon") +
  theme_void()
```

Not the ideal theme for this plot, since removing the axes and their labels leaves us with little information on what's being plotted. But, we know we have the option to change the theme!

A list of themes can be found in the `ggplot2` documentation [here](https://ggplot2.tidyverse.org/reference/ggtheme.html).

## Handy Layers

We were introduced to geoms, labels and themes in the previous section. 

But there are a number of other layers you might find handy for making plots and figures with `ggplot()`:

#### Labels

We saw earlier that we could specify labels using `labs()`, like in Example 1: 

```{r ggplot-labels1}
mtcars %>% ggplot(aes(x = wt, y = mpg)) +
  geom_point() +
  labs(title = "Miles vs. Weight", 
       x = "Car Weight (1000 lbs)", 
       y = "Miles per US Gallon")
```

But we can also make the *exact* same plot by **specifying labels separately**, in their own layers, with `ggtitle()`, `xlab()`, and `ylab()`:

```{r ggplot-labels2}
mtcars %>% ggplot(aes(x = wt, y = mpg)) +
  geom_point() +
  ggtitle("Miles vs. Weight") + 
  xlab("Car Weight (1000 lbs)") +
  ylab("Miles per US Gallon")
```

#### Adjusting title positions using the **theme** layer

Notice from our Labels section examples that by default, `ggplot()` will position our titles to the left.

We can position the title to the centre using the theme layer: `theme(plot.title = element_text(hjust=0.5))`:

```{r ggplot-titlecentre}
mtcars %>% ggplot(aes(x = wt, y = mpg)) +
  geom_point() +
  labs(title = "Miles vs. Weight", 
       x = "Car Weight (1000 lbs)", 
       y = "Miles per US Gallon") +
  theme(plot.title = element_text(hjust=0.5))
```

Similarly, we can adjust the title to the left with `hjust=0` (the default), or to the right with `hjust=1`:

##### To the left, to the left...

```{r ggplot-titleleft}
mtcars %>% ggplot(aes(x = wt, y = mpg)) +
  geom_point() +
  labs(title = "Miles vs. Weight", 
       x = "Car Weight (1000 lbs)", 
       y = "Miles per US Gallon") +
  theme(plot.title = element_text(hjust=0))
```

##### Writing to the right-ing

```{r ggplot-titleright}
mtcars %>% ggplot(aes(x = wt, y = mpg)) +
  geom_point() +
  labs(title = "Miles vs. Weight", 
       x = "Car Weight (1000 lbs)", 
       y = "Miles per US Gallon") +
  theme(plot.title = element_text(hjust=1))
```

#### Axis scale limits

Notice how in our previous plots, the y-axis started at 10. 

But this doesn't give us an accurately scaled representation of the relationship between the x-axis and y-axis variables, so we may prefer to start the y-axis at 0 instead.

We can change the axis scale limits using the `xlim()` and `ylim()` layers.

Here, we set the x-axis scale to range from 0 to 6, and the y-axis scale to range from 0 to 35:

```{r ggplot-scalelim}
mtcars %>% ggplot(aes(x = wt, y = mpg)) +
  geom_point() +
  labs(title = "Miles vs. Weight", 
       x = "Car Weight (1000 lbs)", 
       y = "Miles per US Gallon") +
  xlim(0, 6) + 
  ylim(0, 35)
```

#### Axis text angle

Suppose we wanted to change the **orientation** of our axis labels' scales. 

We can do so with the `theme()` layer.

For example, we can rotate the y-axis scale labels by 90 degrees as so:

```{r ggplot-textangle}
mtcars %>% ggplot(aes(x = wt, y = mpg)) +
  geom_point() +
  labs(title = "Miles vs. Weight", 
       x = "Car Weight (1000 lbs)", 
       y = "Miles per US Gallon") +
  theme(axis.text.y = element_text(angle = 90))
```

Notice that this only changes the angle of the axis *scale* labels, and not the labels themselves!

We can similarly change the x-axis scale label using `axis.text.x` in place of `axis.text.y`.

## Additional Resources and References

* [`ggplot2` cheatsheet for quick reference](https://rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf)

* [`ggplot2` complete reference](https://ggplot2.tidyverse.org/reference/)

* [`ggplot2` themes reference](https://ggplot2.tidyverse.org/reference/ggtheme.html)

* [`ggplot2` title positioning](https://rpubs.com/tuyenhavan/282317)

* [`ggpubr` package reference, which can be helpful for arranging plots in a grid](https://cran.r-project.org/web/packages/ggpubr/ggpubr.pdf)



## Exercises

### Question 1

### Question 2

### Question 3

### Question 4

### Question 5

### Question 6

### Question 7

### Question 8

### Question 9

### Question 10
