
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(flair)
library(knitr)
```

# Scatter plots

*Written by Haoluan Chen and last updated on 7 October 2021.*


## Introduction

In this lesson, you will learn how to:

- Create a scatter plot in R using ggplot2 package
- Customize your scatter plot

Prerequisite skills include:

- Install and load ggplot2 packages

- Basics of using ggplot2: e.g. `ggplot()`, `aes()`.

Highlights:

- Create and customize a scatter-plot


## The content

## What is a Scatter plot?

A scatter plot is a visual representation of two numerical variables. It allows you to see the correlation between the variables. 

Here, we have a dataset `mtcars` extracted from the 1974 Motor Trend US magazine.

```{r scatter1}
head(mtcars)
```

`wt` is the weight (1000lbs), `mpg` is the miles per gallon for the car. Now we are interested in the relationship between these variables among 32 different cars.

Using the following code, we can create our basic scatter plot in R!

```{r scatter2}
# Basic scatter plot
ggplot(data = mtcars, mapping = aes(x=wt, y=mpg)) +
  geom_point()
```

In this plot, each dot shows one car's miles per gallon versus its weight. We can see a decreasing trend among the weight and the miles per gallon for the car.


## Arguments

### Basic scatter plot
- The size and the shape of points can be changed using the function geom_point() as follow :

geom_point(size, shape)

```{r scatter3}
# Change the point size
ggplot(mtcars, aes(x=wt, y=mpg)) +
  geom_point(size=2)
# Change the point shape
ggplot(mtcars, aes(x=wt, y=mpg)) +
  geom_point(shape=23)
```

From the above figures, we know that we can change different sizes and shapes by changing parameters in `geom_point()`.

- The size of the points can be controlled by the values of a continuous variable as in the example below.


`cyl` is the number of cylinders of the car.

```{r scatter4}
# Change the point size by the values of cyl
ggplot(mtcars, aes(x=wt, y=mpg)) + 
  geom_point(aes(size=cyl))
```

## Change the appearance of the point
We can also change the color of the border of the point and filling of the point using the function geom_point() as follow :

geom_point(color, fill)

```{r scatter5}
# change shape, color, size
ggplot(mtcars, aes(x=wt, y=mpg)) +
  geom_point(shape=23, color="darkred",size=3)

# change shape, color, fill, size
ggplot(mtcars, aes(x=wt, y=mpg)) +
  geom_point(shape=23, fill="blue", size=3)

```


## Scatter plots with multiple groups

We can also change the points, colors, and shapes based on a variable.
In the R code below, point shapes, colors, and sizes are controlled by the levels of the factor variable cyl :

**We are using as.factor here because cyl is a double data type in the dataset. Using as.factor() can transform cyl into factor data type**

```{r scatter6}
# Change point shapes by the levels of cyl
ggplot(mtcars, aes(x=wt, y=mpg)) +
  geom_point(aes(shape=as.factor(cyl)))
# Change point shapes and colors
ggplot(mtcars, aes(x=wt, y=mpg)) +
  geom_point(aes(shape=as.factor(cyl), color=cyl))
# change point shapes,  colors and sizes
ggplot(mtcars, aes(x=wt, y=mpg)) +
  geom_point(aes(shape=as.factor(cyl), color=cyl, size=cyl))
```




## Exercises

## Exercise 1 

Create a scatter plot to see the correlation between the hp and wt in mtcars

```{r scatter7}
mtcars
```

```{r scatterplotex1, exercise=TRUE, exercise.lines = 5}

```

```{r scatterplotex1-solution}
mtcars %>% ggplot(aes(hp, wt)) + geom_point()
```

## Exercise 2
Set the size of the scatter plot in previous question to 5 

```{r scatterplotex2, exercise=TRUE, exercise.lines = 5}

```

```{r scatterplotex2-solution}
mtcars %>% ggplot(aes(hp, wt)) + geom_point(size = 5)
```
## Exercise 3
Set the color of the scatter plot in previous question to blue

```{r scatterplotex3, exercise=TRUE, exercise.lines = 5}

```

```{r scatterplotex3-solution}
mtcars %>% ggplot(aes(hp, wt)) + geom_point(size = 5, col = "blue")
```

## Common Mistakes & Errors

- If you want to change the colors and shapes based on a variable in your dataset, make sure it is a factor type dataset
- connect ggplot() and geom_point() by using `+` not `%>%`


## Next Steps

You may use geom_smooth() function to add a regression line on your scatter plot to get a clear pattern of your data. 

```{r scatternext}
ggplot(data = mtcars, mapping = aes(x=wt, y=mpg)) +
  geom_point() + 
  geom_smooth(method='lm') 
```


Here is a guide for creating interesting scatter plots! http://www.sthda.com/english/wiki/ggplot2-scatter-plots-quick-start-guide-r-software-and-data-visualization


You may also change the title, axis, theme of your scatter plot! 

http://www.sthda.com/english/wiki/ggplot2-title-main-axis-and-legend-titles



## Exercises

### Question 1

### Question 2

### Question 3

### Question 4

### Question 5

### Question 6

### Question 7

### Question 8

### Question 9

### Question 10
