
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# leaflet

*Written by Isaac Ehrlich and last updated on January 31st 2022.*

## Introduction
`leaflet` is an R package that is used to create interactive maps. These are especially useful when your goal is to make user-friendly interactive maps for the web. Install the package and then get going!

Prerequisite Skills:

- Basic working knowledge of R
- Be sure to install leaflet!

## Point Maps

Each map in leaflet begins with the `leaflet()` function, a blank canvas onto which maps and layers of points or other representations of your data can be added. Below is a simple example of an interactive map with a single point.
```{r, warning = FALSE}
library(leaflet)

leaflet() %>% # Create blank canvas
  addTiles() %>% # Add default leaflet map
  addMarkers(lat = 43.66, lng = -79.39, label = "University of Toronto") # Add point location and label
```

Rather than manually entering point data, you can also reference data points from appropriate data sets:
```{r}
# Show 10 largest magnitude earthquakes from R's quakes data set
quakes %>% 
  slice_max(mag, n = 10) %>% 
  leaflet() %>% 
  addTiles() %>% 
  addCircleMarkers(lat = ~lat, lng = ~long, label = ~mag) # Reference columns in the quakes data set with '~'
```

Note that there are many available options to customize your point markers. You can visit [leaflet documentation](https://rstudio.github.io/leaflet/) for more examples and complete options.

## Polygons in Leaflet
Leaflet is also capable of handling more complex data, such as polygons and lines. Polygons can be loaded into leaflet through shapefiles, and will require that both `.shp` and `.shx` files are present in your directory. In the following example, we will use the `rgdal` package to read in shape data.

```{r, warning = FALSE, message = FALSE}
library(rgdal)
ecoregions <- readOGR(dsn = "data", layer = "canada-ecoregions", verbose = FALSE) # Read shapefiles

leaflet(ecoregions) %>% 
  addProviderTiles(providers$Stamen.Toner) %>% # Add custom base map
  addPolygons(color = "#444444", weight = 1, smoothFactor = 0.5, 
    opacity = 1.0, fillOpacity = 0.5, fillColor = "#E39B3D") # Add polygons from shapefile
```


## Questions
1.
Leaflet maps are interactive.
  a. True
  b. False

2.
The function `leaflet()` never takes any arguments and is just used as syntax sugar to indicate the type of map.
  a. True
  b. False

3.
Which of the following is the correct special character used when adding markers?
  a. `addCircleMarkers(lat = ~my_lat, lng = ~my_long)`
  b. `addCircleMarkers(lat = }my_lat, lng = }my_long)`
  c. `addCircleMarkers(lat = %my_lat, lng = %my_long)`
  d. `addCircleMarkers(lat = ;my_lat, lng = ;my_long)`
  
4.
Leaflet is a very limited package. It can only display points on a map.
  a. True
  b. False

5.
How can you read shape data into R?
  a. There is no way to read shapefiles, you need to convert them to another file type externally first
  b. Using `read.shp()`
  c. Using `readOGR()` from the rgdal package
  
### Question 6

### Question 7

### Question 8

### Question 9

### Question 10
