
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# Stan

*Written by Marija Pejcinovska and last updated on 6 February 2022.*


## Introduction

In this tutorial we will cover the basics of Stan and how to use it with the help of R. Stan is a probabilistic programming language (more on this in a bit) that is used in Bayesian analysis. For the purposes of this tutorial, we will assume that you are already familiar with the Bayesian framework of inference and we will, therefore, not cover it. The tutorial is largely an introduction to the building blocks of a Stan program and is designed to help you get started on using Stan for your Bayesian modeling tasks. 


In this lesson, you will:

- Learn what  Stan is.
- Learn some of the Stan syntax and the basics of a Stan program.
- Learn how to install and get started with RStan and CmdStanR, both of which are R interfaces to Stan.

Prerequisite skills include:

- Basic familiarity with Bayesian analysis and MCMC sampling. 
- The module on installing packages.
- The modules on asking for help. 


## What is Stan and why should we care about it?

Stan is a probabilistic programming language (PPL) designed to automate Bayesian inference. What this means is that Stan is not a full fledged language the way traditional programming languages are (think here of C++, Python, Haskell, Java and so on), but is a program (one that is specifically written in C++) equipped with random generators and other functions and routines suitable for carrying out Bayesian analysis. Stan is not the only PPL out there, but is one that has recently been used heavily by statisticians and practitioners of statistics. You may have also heard of JAGS or BUGS — another pair of PPLs frequently used in statistics.  
What is different about Stan is what's under the hood. Stan generates representative samples from a posterior distribution using a Hamiltonian Monte Carlo (HMC) algorithm, or specifically a variation of it know as No-U-Turn Sampler, or NUTS for short. This makes Stan faster and more efficient than JAGS and BUGS (which use a set of different samplers to generate the Monte Carlo steps), especially for very large complex models. Of course, Stan in not uniformly better. There are a number of things that Stan is not well equipped to do. One important one is that the sampler in Stan (HMC) only allows for parameters defined on a continuous space, meaning that, depending on your model or analysis goal, dealing with discrete parameters may be difficult or altogether not possible to do in Stan.


#### And.. why the name Stan?
If you were curious about the name, it turns out that it's a nod to Stanislaw Ulam — a Polish mathematician and a physicist who was one of the pioneers of the Monte Carlo methods. 
If you are a nostalgic elder Millennial or a Gen Z kid something entirely different might come to mind when you hear the word "Stan". But if you are also an enthusiastic Bayesian, you might just totally stan for Stan! (Side note: This is the only terrible joke in the whole tutorial; I promise!)

## Installing R interfaces to Stan

**Before proceeding with any installations, make sure that you are using the most current versions of R and RStudio!**  
  
The Stan modeling language and its algorithms can be accessed through a number of interfaces to many of the popular computing languages and software (R, Python, Scala, MATLAB, Stata). This toolkit is devoted to R, so we will focus on two main R interfaces to Stan, RStan and CmdStanR. We recommend installing and using CmdStanR, the newer of the two packages, as it is lighter weight and installation has proven to be smoother. However, you may have your own reasons for installing RStan, and we have included links to those installation resources as well. 

Because installation instructions have been changing rapidly and are somewhat machine-specific, in this tutorial we will only share a few notes on installation and mainly point you in the direction of installation guides set up by the RStan and CmdStanR's development teams.

**NOTE**: If you are using a Mac machine with the latest M1 chip you might want to work with CmdStanR for now (now being late 2021/early 2022 — this will probably change in time)! If during the installation process you encounter problems configuring the C++ toolchain, [this Twitter thread](https://twitter.com/rlmcelreath/status/1479484823991472134) may have helpful advice. 

In general, if you are struggling with installation, remember that Stack Overflow and Stack Exchange are your friends and folks on those platforms might be able to help. You can also ask for help on the Stan Forum (https://discourse.mc-stan.org/).

### Installing RStan
RStan was the first R interface to Stan to be developed. The following homepage should be your starting point: https://mc-stan.org/users/interfaces/rstan.html. The [Getting Started Guide](https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started) outlines the steps necessary to install RStan.   

  - Make sure you verify that your R installation can compile C++ code (follow the steps on configuring C++ Toolchain). Please note that depending on you operating system this may be a non-trivial task and it is hard to predict what type of error you may encounter (if any). 
  - Use the code provided in the Quick Start Guide to proceed with installing and verifying RStan.


### Installing CmdStanR

CmdStanR is a lightweight R interface to Stan and was built as an alternative to RStan. It is technically a wrapper for CmdStan, which itself is a shell interface to Stan.  
To get started with installation of CmdStanR and learn some more about it use the following link: 
https://mc-stan.org/cmdstanr/  

  - Note that CmdStanR requires a working version of CmdStan, which CmdStanR can install for you.  
  - Make sure you have a suitably configured C++ toolchain (instructions on this can be found at the link).


## The building blocks of a Stan program 

Now that you have (hopefully!) successfully installed CmdStanR (or RStan) we can start discussing some concepts that should help get you more familiar with aspects of a Stan program. But before we get there, let's see what a Stan program actually looks like.


### A quick example program to start us off - Biased coin?

[Based on a simple example provided in Carpenter et al., 2017 — see the Additional Resources section]  

Suppose we are interested in learning something about the unknown probability of heads from 10 coin tosses. In this case we'll represent our data by $y$, where each $y_i \in \{0,1\}$, heads (our successes) are represented by 1 and tails by 0 and coin tosses are assumed independent. Then $y_i$ follows a Bernoulli distribution with some probability of "success" $p$. Suppose we assign a uniform prior to $p$, which in this case, since $p$ is defined on the interval $[0,1]$, is equivalent to a $Beta(1,1)$ distribution.  
Imagine that we observe the following data:
```{r}
y <- c(1,1,0,0,0,1,0,1,1,1) # our actual sample from 10 trials
```
Then a Stan model can be specified as follows:
```{r, echo=TRUE, eval=FALSE}
# as string
coin_model_in_Stan = "

data{
  int<lower=0> N; // number of observation
  int y[N]; // vector of length N with our binary observations
  // could also specify as: int<lower=0, uppper=1> y[N];
}

parameters{
  real<lower=0, upper=1> p; // the probability of heads, p
}

model{
  p ~ beta(1,1); // the prior for p
  y ~ bernoulli(p); // the model for the data, i.e the likelihood
}

"

```

To make it easier, we have defined the model here as a string, however, you can specify and save your model as a separate `.stan` file (`.stan` is one of the options for a new file type in RStudio — check it out! it comes equipped with a basic model script). 

## More on the building blocks of a Stan program 

### Data types 
Because Stan relies on a statically-typed language (it's written in C++, which is one such language) all types must be explicitly declared by the user at compile-time. In our example program you may have noticed declaration statements before each one of our variables (for example declaring type`int` for `N` and `y`). And so in Stan you need to declare the type (and possibly constraints) of all your variables before you proceed with compiling your program; once you declare each variable type, it remains unchanged throughout the execution of the program. 

The **primitive types** in Stan are `real` and `int`. `real` is used for continuous value, whereas `int` is for integer values. Integer expression can be used anywhere a real value is required, but not the other way around.     

Stan provides three **matrix-based data types**; `vector` for column vectors, `row_vector` for row vectors, and `matrix` for matrices. Vectors are declared with their sizes and matrices with the number of their rows and columns, and are all accessed using the usual bracket notation; for example, `x[2]` is the 2nd element of the vector `x` and `Z[1,2]` is the element in the 1st row and 2nd column of the `Z` matrix. Note that indexing is Stan begins at 1.     

Stan also supports **arrays**. Any of the data types discussed above can be turned into an array type by using the array argument. For instance, `array[2,2] int b;` declares `b` to be a two-dimensional array of integers with 2 rows and 2 columns, and `array[4] vector[5] y;` declares `y` to be an array of size 4 containing vectors of length 5.  


Stan allows you to declare **variables with constraints**. The constraints might have different effect depending on the block where the variable is defined (more on program blocks in a bit). All of the data types we have discussed can be declared with a lower bound, an upper bound, or both, using relatively standard syntax such as: `int<lower = 1> N;` or `vector<lower = 0, upper = 1>[5] p;`. Stan also provides special data types for structured vectors and matrices. For more on these consider consulting the [Stan Reference Manual](https://mc-stan.org/docs/2_28/reference-manual/index.html).  

**NOTE**: In Stan, the *size* declared for each variable as well as any *constraints* associated with a variable are *not* part of a variable's underlying data type. This means that they are determined dynamically (at run-time) and are not part of the check that Stan does at compile-time.


### A note on Stan syntax and specifications

If you look carefully at our example model you will notice that there are three distinct blocks — a *data block*, a *parameters block*, and a *model block*. **Every Stan program is organized into a sequence of named blocks**. Ours is just a simple model, so using just these three blocks is sufficient; however, Stan allows for much more complex models which may require calling on additional components of the program (more on Stan's program blocks in the next section).   

You will also notice from the model above that **each expression and statement within each block ends with a semi-colon** (**`;`**). This is the end-of-command marker used in C++, which, as we've mentioned before, is the underlying language in Stan. Missing semi-colons are a frequent reason why your Stan program may not compile (Stan beginners accustomed to R's use of `return` may struggle a bit with this in their early Stan use).   

**Comments** in Stan are indicated by `//`; slightly different from R's use of `#`.  

*A note on naming variables*: Variables in Stan should be strings containing only lower-case and upper-case letters, digits, and the underscore (**`_`**) character. No other special characters are allowed. Variables must always start with a letter (either lower- or upper-case), and they should never end with two underscore characters (**`__`**). There are certain names that are considered *reserved* in Stan. Refer to the [Manual](https://mc-stan.org/docs/2_28/reference-manual/index.html) for more details on this. 

Stan encourages **vectorization**. For instance, the vectorized likelihood in the model above ` y ~ bernoulli(p);` can also be coded using a loop: 
```{r, eval=FALSE}
for(n in 1:N){
  y[n] ~ bernoulli(p);
}
```
However, this will run more slowly than the vectorized form!

## Yet more on the building blocks of a Stan program 

### General structure of a Stan model

There are seven named blocks in a Stan model program. An outline of a Stan program with all its blocks is given below


```{r, eval = FALSE}

functions {
  // ... relevant function definitions ...
}

data{
  // ... some experssions/declarations ... 
}

transformed data {
  // ... some experssions/declarations ... 
  // ... some statements ... 
}

parameters{
  // ... some experssions/declarations ... 
}

transformed parameters {
  // ... some experssions/declarations ... 
  // ... some statements ... 
}

model {
  // ... some experssions/declarations ... 
  // ... some statements ... 
}

generated quantities {
  // ... some experssions/declarations ... 
  // ... some statements ... 
}


```

All of the blocks in a Stan program are technically optional. While you may pass an empty string as a valid Stan program, you should expect to encounter errors at the compiling stage. One important thing to keep in mind is that the blocks of your Stan program (even if you only use a few as we did in our basic example) must be written in the order shown above.   

Below we look at some of the blocks in a bit more detail.  

**`functions` block**: This block is reserved for any special, user-defined functions that are to be used later on in the program.  

**`data` block**: The data block is where you declare as variables the data that is read in and used to fit the model. No transformations of the data take place in this block.  

**`transformed data` block**: This block can be used to define new variables that can be computed based on the data defined in the data block. In this block you should declare the desired variable and follow that declaration by a statement (or sequence of statements) that defines that variable. Since statements in the transformed block are designed to be executed once and have a deterministic result, no sampling statements (such as `x ~ normal(0,2)`) should be used in this block.

**`parameters` block**: The variables declared in the parameters block correspond to the variables being sampled by Stan. In our simple example of coin tosses, **`p`**, the probability of heads, is a parameter. Note that in this block, variables declared as parameters can't be assigned any values. 


**`transformed parameters` block**: Here parameters declared in the parameters block can be transformed to other parameters using declarations and statements. Notice that this is quite analogues to what is done in the transformed data block. For instance, suppose your parameter block defines the following parameters
```{r, eval=FALSE}
parameter {
  real<lower=0, upper=1> lambda;  // some prior parameter
  real<lower=1> beta;             // another prior paramter
}
```
Then your transformed parameters block might look something like
```{r, eval=FALSE}
transformed parameters {
  real<lower=0> gamma = lambda*beta; 
}

```


**`model` block**: The statements in the model block define the model, or more specifically the log densities on the constrained parameter space. This block typically includes sampling statements. For instance, in our coin toss example we have two simple sampling statement; the parameter $p$ has a uniform prior distribution and each toss $y[i]$ is drawn from a Bernoulli($p$). The model block is executed after the transformed parameters block each time the log density is evaluated.  

**`generated quantities` block**: This is where you can efficiently define any values that depend on the model parameters and data, but do not affect estimation. The generated quantities block is a bit different than the rest. The block is executed only after the sample has been generated and nothing in this block affects the sampled values. It is generally recommended that a quantity that does not play a role in the model be defined in this block. Note also that all variables declared as generated quantities are printed as part of the output. This a good place to code predictive values, transform parameters for reporting, calculate log-likelihoods, deviances and various other measures used in model comparison, and so on.


### A few additional notes on program blocks and specification in Stan

  - The `data` and `parameters` blocks are the only blocks that do not allow for statements of any kind. As briefly mentioned before, this is because the data and parameters blocks are used exclusively to declare the data variables that are inputs for the model and the parameter variables that are to be sampled.   
  - It's good to keep in mind that the lines in a Stan model specification are processed in the order in which they are written. Variables must always be declared before they are used in a statement. In fact, in Stan we describe models *top-down*. This means that we start with the hyperpriors (if any), then the priors, and in the end we introduce the likelihood. 
  



## Compiling a Stan program and getting posterior samples

### Revisiting our coin example

Recall our coin toss example.   
We had the following data on 10 coin tosses
```{r}
N <- 10 # number of trials
y <- c(1,1,0,0,0,1,0,1,1,1) # our actual sample of 10 coin tosses
```

and specified the following Stan model to make inference about the probability of success (heads), `p`:
```{r, echo=TRUE, eval=TRUE}
# as string
coin_model_in_Stan = "
data{
  int<lower=0> N; // number of observation
  int y[N]; // vector of length N with our binary observations
  // could also specify as: int<lower=0, uppper=1> y[N];
}

parameters{
  real<lower=0, upper=1> p; // the probability of heads, p
}

model{
  p ~ beta(1,1); // the prior for p
  y ~ bernoulli(p); // the model for the data, i.e the likelihood
}

"

```

### Translating and Compiling

Once you have correctly specified the model in Stan, the next task is to translate it into a C++ code and compile it into an object (specifically, an executable *dynamic shared object*) which can then be used to generate the Monte Carlo samples from the posterior distribution. This may sound complicated, but is in fact quite easy to do in either RStan or CmdStanR. Here we'll demonstrate this using appropriate commands from CmdStanR; however it is equally easy to do this using the RStan library.  

We'll use the command `cmdstan_model` to let Stan translate and compile the model. Recall that in the compile stage we are just readying the program to generate posterior samples, but we have not generated those samples yet. Before you do anything, make sure that your `cmdstanr` library is loaded 

Because we defined the model as string, we'll additionally need to use the `write_stan_file()` command.

```{r, eval=TRUE, warning=FALSE, message=FALSE}
library(cmdstanr) 

coin_mod <- cmdstan_model(write_stan_file(coin_model_in_Stan))

# if you have your model program saved as a .stan file, say coin.stan, you can just use 
# coin_mod <- cmdstan_model("coin.stan") # make sure that the file paths to your stan file are correct
```
This translation and compilation step can take a bit of time, particularly if your model is quite complex. This is also where Stan will alert you to any syntax errors in your program.   

The object `coin_mod` is now an object of class CmdStanModel. In CmdStanR any relevant methods for this object are accessed using the `$` operator.  

### Getting MCMC samples 

The `$sample()` method for CmdStanModel objects runs Stan’s default MCMC algorithm. This will allow us to get our posterior samples. 
First, however, we need to specify the data. The `data` argument in the `$sample` method accepts a named list of R objects (same as in RStan). Make sure that the names of the objects you pass in the data match those defined in the `data` block in your Stan model. 

```{r, eval=TRUE, echo=TRUE}
coin_data <- list(N = 10, y = c(0,1,0,0,0,0,0,0,0,1))

coin_fit <- coin_mod$sample(
  data = coin_data, # our data passed as a named list
  seed = 123, 
  chains = 1, # number of chains (here chosen to be 1 for simplicity)
  iter_warmup = 1000, # number of warmup iterations to run per chain
  iter_sampling = 1000 # number of post-warmup iterations to run per chain
)
```

`coin_fit` is now a CmdStanMCMC object. There are a umber of methods you could use to access aspect of this object.  

To run posterior summary statistics, use the `$summary()` method which calls functions from the `posterior` package (make sure you have the package installed and loaded before doing this!!).  

```{r, eval=TRUE, message=FALSE, warning=FALSE}
library(posterior) 
coin_fit$summary()
```

To extract posterior draws the `$draws()` method can be used. Posterior draws can be extracted in a number of formats provided by the `posterior` package. Here we'll highlight one particular format, and that is `draws_df`, where the output is a dataframe with dimensions determined by the draws (rows) and variables outputted (columns). Here the draws represent iterations*chains, and the output variables are your model parameters as well as `lp__` (the log density (up to a constant)); there are also some other (hidden) elements of the MCMC samples. 

```{r, eval=TRUE}
coin_draws_df <- coin_fit$draws(format = "df")
coin_draws_df
str(coin_draws_df)
```

## Where to next?

This tutorial was designed to give you a very brief introduction to Stan, walk you through the basic components of a Stan program, and get you started with CmdStanR. Your Bayesian  analysis tasks, however, are likely to require a bit more than what was covered so far.  
You may want to consider looking into available CmdStanR vignettes for more on the various features of CmdStanR, as well as some additional functionalities of the `posterior` package. 

For visual summaries of your findings consider the package `bayseplot`. And for model diagnostics and model comparisons you might want to get familiar with the capabilities of the `loo` package. 

Of course, **you don't need to write your own Stan programs to carry out Bayesian analysis**. Stan can certainly seem overwhelming at first, and there isn't always a need to code everything up from scratch! The R interfaces to Stan which were introduced in this tutorial are considered lower-level and that is a bit of a downside. There are fortunately some user-friendly higher-level R interfaces to Stan. The packages `RStanArm` and `brms` are such examples. Bear in mind that `RStanArm` requires RStan in the background. `brms`, on the other hand, seems to support CmdStanR as an alternative backend to RStan. Both use similar syntax as the usual modeling packages you may be familiar with, such as `lm`, `glm`, `lme4` and so on. Both packages provide excellent infrastructure that gives you a lot of wiggle room to run relatively complicated Bayesian models in a very automated and intuitive way.


## Additional resources 

 - For more technical details, check out the original Stan paper *"Stan: A Probabilistic Programming Language"* by Carpenter et al. published in the Journal of Statistical Software (Vol 76, Issue 1, 2017), which you can find [here](https://www.jstatsoft.org/article/view/v076i01)    
  
  - All the documentation about Stan can be found at: https://mc-stan.org/users/documentation/. This includes a Stan User's Guide, a Stan Reference Manual, and a Functions Reference. This is also where you'll find documentation on CmdStan, CmdstanR, and RStan, as well as ample Stan resources, including various case studies and tutorials on Stan. 
  
  - If you are interested in learning more about Hamiltonian Monte Carlo (HMC), the sampling algorithm in Stan, you can check out Michael Betancourt's paper on HMC, ["A Conceptual introduction to Hamiltonian Monte Carlo"](https://arxiv.org/pdf/1701.02434.pdf), available on arXiv. 



<!--
## Exercises


#### **Exercise 1**  

Suppose for a given problem we have observations on `N` individuals and we have some covariate `x` that we are interested in using in our regression model. Take a look at the following statement and answer the questions below.
```{r , eval = FALSE}

vector[N] x_stand = (x - mean(x))/ sd(x);

```

```{r ex1q1, echo = FALSE}
question("The expression above is syntactically correct?",
         answer("TRUE", correct = TRUE),
         answer("FALSE"),
         allow_retry = T)
```

```{r ex1q2, echo = FALSE}
question("In which of Stan's named blocks does the expression above belong?",
         answer("functions"),
         answer("data"),
         answer("transformed data", correct = TRUE),
         answer("transformed paramaters"),
         allow_retry = T)
```


#### **Exercise 2**  

Consider the following expression
```{r, eval = FALSE}
real new_sumlog(int N, vector x, real k){
   vector[N] p;
   real lp;
   for (i in 1:N){
     p[i] = k*exp(-k*x[i]);
   }
   lp = sum(log(p));
   return lp;
 }
```

```{r ex2q1, echo = FALSE}
question("In which of Stan's named blocks does the above expression belong?",
         answer("functions", correct = TRUE),
         answer("data"),
         answer("parameters"),
         answer("transformed paramaters"),
         allow_retry = T)
```

-->

## Exercises


### Question 1


Suppose for a given problem we have observations on `N` individuals and we have some covariate `x` that we are interested in using in our regression model. Consider the following expression:  

```{r , eval = FALSE}

vector[N] x_stand = (x - mean(x))/ sd(x);

```

In the syntax of Stan this is a correctly specified statment.  

  a.  True  
  b. False

### Question 2

Consider once more the expression in Question 1. In which of Stan's named block does the expression belong?  

  a. The `functions` block.  
  b. The `data` block.  
  c.  The `transformed data` block.  
  d. The `transformed parameters` block.  
  
### Question 3

In the Stan language, the data types `int` and `real` can be applied interchangeably.  

  a. True  
  b.  False  
  
### Question 4

Which of the lines below give the correct syntax to represent an object `M` which is an array of size `N` containing matrices of size `l` by `k`.  

  a.  array[N] matrix[l,k] M;  
  b. array M matrix[l,k];  
  c. matrix[l,k] M[N];

### Question 5  

Just like in R, a `;` to mark the end of a command in Stan is technically optional.  

  a. True  
  b.  False  
  
  
### Question 6

Suppose one of our covariates in a model, call it `x_extra`, is generated randomly from a Normal distribution as `x_extra ~ normal(0,3)`. We can place this expression in the `data` block.  

  a. True  
  b.  False  
  
### Question 7  

Consider the following expression
```{r, eval = FALSE}
real new_sumlog(int N, vector x, real k){
   vector[N] p;
   real lp;
   for (i in 1:N){
     p[i] = k*exp(-k*x[i]);
   }
   lp = sum(log(p));
   return lp;
 }
```

In which one of Stan's named block can we write this expression? (Hint: Think about what type of expression this reminds you of)  

  a. The `data` block.  
  b.  The `functions` block.  
  c. The `parameters` block.  
  d. The `transformed parameters` block.

### Question 8 

Suppose in a Stan program we have correctly specified two covariates `X` and `W` as `matrix[N, M] X;` and `matrix[N, L] W;`. An additional covariate `Z` specified with the statement `matrix[N, M + L] Z = append_col(X,W)` should be placed in which of Stan's named blocks?  

  a. The `data` block. 
  b.  The `transformed data` block.
  c. The `functions` block. 
  d. The `generated quantities` block.  
  

### Question 9  

The data used in a Stan model is supplied at compile time.  

  a. True  
  b.  False

### Question 10 

Stan generates posterior samples using which sampler?  

  a. A fancy Metropolis-Hastings algorithm. 
  b. A Gibbs sampler.  
  c.  A Hamiltonian Monte Carlo algorithm.
