
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(sortable)
library(testthat)
```

# testthat

*Written by Matthew Wankiewicz and last updated on January 5th, 2022*

## Introduction

Once you have completed making an R package, or while you are making it, you may want to test your functions to make sure they are working properly. These tests can be as simple as running the function in the console and seeing if the result matches what you expect. The `testthat` package allows us to run these tests automatically and keeps them located in a separate folder.

This package is very useful for testing as it can test functions for many different cases, including if a function will throw an error, or if a function's output matches what you expect it to be. 

On the package's website, it says "testthat tries to make testing as fun as possible, so that you get a visceral satisfaction from writing tests. Testing should be addictive, so you do it all the time". `testthat` achieves this by being easy to use, and very easy to understand.  

## Loading in the Package

`testthat` is located on CRAN and can be installed using `install.packages("testthat")`. To use `testthat` with your package, you will need to run `usethis::use_testthat`. This will create a folder called "testthat" where you can keep your tests. Along with creating a testing folder, the function will also add a file that will run all of your package's tests. 

## Setting up your tests

After you have loaded `testthat` into your session, you can begin to write tests for your functions. The tests can cover a wide variety of cases. The most common testing functions to use are `expect_` functions. These functions test certain cases in your package's functions, such as if there will be an error or if the output is what you expect it to be. 

To use the functions, you will create files inside of the "testthat" folder for the functions you want to test. Next, make sure to name the file so it matches the function you plan to test. Your file should be named "test_functionname". This is important because if any issues appear while running the tests, `testthat` will use the file name to tell you where the issues occur. 

## `expect_` functions

This section will look at the most common `expect_` functions and what they do or how they can be used.

#### - `expect_equal` & `expect_identical`

These two functions are very similar as they look at your functions output and checks if the output is the same as what you expect it to be.
The main arguments for these functions are:

* `object`: This is the function case you want to test. For example, you could set this as 1+1
* `expected`: This is what you expect the output of the function to be. For the example above, this would be set to 2.

The chunk below shows an example of the `expect_equal` function, where the object does not match the expectation.
```{r expect_equal-example, error=T}
expect_equal(2*2, 2)
```

Since 2*2 does not equal 2, the function give an error and tells us what the issue is.

When we run the function using an object that matches the expectation, the function will not return anything.
```{r expect_equal-example2}
expect_equal(nchar("abc"), 3)
```

The `expect_identical` function works in the same way, but is better suited for determining if strings or vectors are the same, instead of numbers.

#### - `expect_type` & `expect_s3_class`

These functions looks at the class of a function output. `expect_type` is meant to be used with more basic operations such as testing if a vector's class is character, integer or numeric. On the other hand, `expect_s3_class` analyzes "s3 classes" which are more advanced classes in R, such as tibbles or data frames. If you try to use `expect_type` on a data frame, the function will say that the data frame has type 'list'. More information about s3 classes can be found at https://adv-r.hadley.nz/s3.html.

If you have a function which changes the class of objects from integer to character, or vice versa, `expect_type` will be useful. If you want to test for the presence of tibbles or data frames, `expect_s3_class` will work best.

Similar to `expect_equal`, these functions take in an object and what you expect the output of the object to be. This time instead of expecting a value, you will enter in a class:

* `object`: The object or function you want to test
* `class`: The expected class of the object

To show how `expect_type` works, we will use the `as.double` function which turns characters into doubles. 

When the cases do not match, the function's error will tell us what the difference is.
```{r expect_type-ex1, error=TRUE}
expect_type(as.double("3.14"), "numeric")
```

If they do match, there is no output.
```{r expect_type-ex2, error=TRUE}
expect_type(as.double("3.14"), "double")
```

To test expect_s3_class, we can make a tibble and check that it's class is "tbl_df".
```{r expect_class-ex1, error=TRUE}
my_df <- tibble("time" = c(1:3), "date" = c("a", "b", "c"))
expect_s3_class(my_df, "tbl_df")
```

If the code above did not make a tibble, we would receive an error message telling us the class of the object.

```{r expect_class-ex2, error=TRUE}
my_df <- tibble("time" = c(1:3), "date" = c("a", "b", "c"))
expect_s3_class(my_df, "integer")
```

We can also use the `$` command to test the class of the rows of data frames and tibbles.

```{r expect_class-ex3, error=TRUE}
my_df <- tibble("time" = c(1:3), "date" = c("a", "b", "c"))
expect_type(my_df$time, "integer")
```

#### - `expect_length`

`expect_length` can be used to test the length of your function's output. This function works best for functions that create vectors of length n, or repeat vectors n times.

The arguments for this function are the object and `n`, the expected length:

* `object`: A function which creates an object of a certain length
* `n`: The expected length of your object

Like the other functions, if there is a mismatch in the object and the expectation, an error will occur, telling us the observed length of the object.

```{r expect_length-ex1, error=TRUE}
expect_length(c(1:5), 2)
```

Again, if there is no difference, there will be no output.

```{r expect_length-ex2, error=TRUE}
expect_length(c(1:5), 5)
```


#### - `expect_lt`, `expect_gt`, `expect_lte`, `expect_gte`

The next group of `expect_` functions look at whether or not the object given is greater/less than the expectation. For these functions, "lt" means less than, and "gt" means greater than. The other two functions, `expect_lte` and `expect_gte`, look at cases of less/greater than or equal to an expected value. 

Since the function specifies if you want to test if the object is less than or greater than a value, the arguments are still an object and an expected value:

* `object`: A value or expression that you want to compare
* `expected`: The value you want to compare to the object

With these functions, if there is a value greater than the object when using the `lt` functions, an error will occur along with the difference between the expected value and the object.

```{r expect_lt-ex1, error=TRUE}
expect_lt(2,1.9)
```

Like the other functions highlighted in this lesson, if the expectation passes, there will be no output or errors given.

```{r expect_lt-ex2, error=TRUE}
expect_lt(2,2.1)
```

#### - `expect_vector`

`expect_vector` can be used to test objects and determines if they are the correct size or match a "prototype" vector. The prototype vector is just the class of the vector (character, integer, double, etc.). This function is similar to `expect_length` and `expect_type` and can be used as a combination of both.

The arguments for this function are:

* `object`: The object or function output you want to examine
* `ptype`: A vector "prototype", this will be the class of the vector. You can use functions like `integer()` or `character()` to test vectors
* `size`: This is the size of the vector

When using this function, it looks for errors in the class of the vector first, and then looks for errors in the size.

If an error occurs with the class or size of the vector, the error message will tell you what the class or size of it is.

```{r expect_vector-ex1, error = TRUE}
expect_vector(c(1:10), ptype = character())
```
Since the classes do not match, an error occurs telling us what the actual class is. The function does the same for the size argument.

```{r expect_vector-ex2, error = TRUE}
expect_vector(c(1:10), size = 9)
```

Here the function is telling us that before it can test the vector, it must have the correct size.

As mentioned earlier, if there are issues with both the size and class of the vector, the error will be for the object class.

```{r expect_vector-ex3, error = TRUE}
expect_vector(c(1:10), size = 9, ptype = character())
```

#### - `expect_error`, `expect_warning`, `expect_message`

This final set of functions looks for errors or warnings. This can be useful for cases where you know errors will occur or expect for them to occur. By using these functions, you can be sure that users are receiving proper instruction when they reach an error or warning and know how to avoid issues in the future.

These functions are different from previous functions as you do not necessarily need to have an expectation in the function call. If your error/warning message is long or difficult to copy, the function will look for an error/warning instead of a specific message. If you do include an expected error/warning message, the function will make sure they match. 

The arguments for these functions are:

* `object`: This is the object you want to test, it will be an error, warning or message given by a function.
* `regexp`: This is the message you expect to receive from the function. This argument is optional and can be set to NA if you expect there to be no issues.
* `ignore.case`: This determines if you want the detection to be case-sensitive. Setting this to "FALSE" makes sure the error messages are identical.

When errors/warnings/messages are found using these functions, there will be no output unless you specify an expectation.

```{r expect_error-ex1, error=TRUE}
expect_error(nchar(), "Missing Argument")
```

Since we did not write the correct message, an error occurs and tells us the correct message.

If we do not put an error message, the function will not return anything.

```{r expect_error-ex2, error=TRUE}
expect_error(nchar())
```

We can also get error messages if there are no issues with the function, telling us that there were no issues.

```{r expect_error-ex3, error=TRUE}
expect_error(nchar("abc"))
```

The same idea applies for the warning and message functions.

### `test_that`

After deciding on which functions you want to use to test out your package. You can place the tests in an R file containing tests for a specific function. The tests in this file can look at many different cases and you can choose to be as specific as you want in your tests. 

To properly use the `expect_` functions, you will write them inside the `test_that` function. This function lets you test multiple cases at once and will alert you of any issues. Similar to the `expect_` functions, if an error occurs, you will see what the issue is and where it occurs. If there aren't any issues, the function will tell you that the tests have passed. 

The arguments for `test_that` are:

* `desc`: This is a description of the cases you want to test. If you plan on testing a function using only odd numbers, your description could "Testing function for odd numbers". The tests can also look for errors or warnings. 
* `code`: This is where the `expect_` functions go. You can place them inside braces "{}" to ensure that the tests are being run properly. You can add multiple function calls in this argument.

To show how the function works, we can test the `nchar` function with words of odd length. We will do this by defining a description and then the test cases.

```{r test_that-ex1}
test_that("nchar function works for words of odd length", {
  expect_equal(nchar("a"), 1)
  expect_equal(nchar("abc"), 3)
  expect_equal(nchar("abcde"), 5)
})
```

Since all of the tests passed, we get a message saying "Test passed". Using a similar set up, we can look at what happens when one of the tests fail.

```{r test_that-ex2, error=TRUE}
test_that("nchar function works for words of odd length", {
  expect_equal(nchar("a"), 1)
  expect_equal(nchar("abc"), 4)
  expect_equal(nchar("abcde"), 5)
})
```

As we see above, we will get a message telling us which line the error occurs on and what the error is. If we have multiple failures in our tests, the function will return all cases where a failure occurs.

```{r test_that-ex3, error=TRUE}
test_that("nchar function works for words of odd length", {
  expect_equal(nchar("a"), 1)
  expect_equal(nchar("abc"), 4)
  expect_equal(nchar("abcde"), 6)
})
```

#### `skip`

If you want to ignore a test, you can use the `skip` function. This will ignore any function calls after `skip` is used. 

The only argument for `skip` is a message explaining why you are skipping the tests. The chunk below shows an example of the skip function being used.

```{r test_that-ex4, error=TRUE}
test_that("nchar function works for words of odd length", {
  expect_equal(nchar("a"), 1)
  expect_equal(nchar("abc"), 3)
  skip("skip")
  expect_equal(nchar("abcde"), 6)
})
```

## Running Tests

After writing your tests and placing them in a testing file, you need to run your tests. To do this, you can go to the window header, select "Build" and then "Test Package". You can also run tests using Control + Shift + T or Command + Shift + T. After using this command, it will run tests for all of your functions. It will split up the test results by each test file. After completing all of the tests, it will tell you how many tests failed, raised warnings, were skipped or passed successfully.

## Common Issues and Errors

- When running tests for your package, you may see that some warnings appear after running the tests. These warnings are usually there to tell you about issues with running functions or pre-set warnings from functions you use. To solve this, you can look at the warning message and see what the issue is. In some cases, you may not need to do anything and can continue working on your package.

- Sometimes when your tests are formatted properly, they can still fail. This happens when you do not fill out a description for your test. When you run your tests without a description, your tests will fail. To fix this, you just need to add a description to the `test_that` function and your tests will run.

- If you see a "argument missing" error, this means that you may have forgotten to add an argument to an `expect_` function, either the object or expectation. To fix this, double check that all arguments are filled out. If that is not the issue, check for typos in the object names. 

## Exercises

### Question 1

```{r testthat-q1, echo=FALSE}
question("True or False: When an object matches the expectation in an `expect_` function, there will be no output from the code?",
         answer("True", correct = T),
         answer("False", message = "When there is a match in these functions, there is no output, an error occurs if there is no match."),
         allow_retry = TRUE)
```

### Question 2

```{r testthat-q2, echo=FALSE}
question("If I want to check that my object (a number), matches my expectation (another number), which function is best suited for this scenario?",
         answer("`expect_equal`", correct = TRUE),
         answer("`expect_identical`", message = "This function is better suited for matching characters as opposed to numbers"),
         answer("`expect_length`"),
         answer("`expect_same`"), 
         allow_retry = TRUE)
```

### Question 3

```{r testthat-q3, echo=FALSE}
question("True or False: You can include multiple test cases in a `test_that` function call?",
         answer("True", correct = TRUE),
         answer("False"),
         allow_retry = TRUE)
```

The next three questions will build upon each other. For the first question, create a `test_that` function with one test for the `expect_lt` function and a proper description. (Note: There will be many correct answers.)

### Question 4
Create a test for the `expect_lt` function that tests if 3.14 is less than a number
```{r testthat-q4, eval=FALSE, exercise=TRUE}
## Enter code below:
```

```{r testthat-q4-solution}
## Enter code below: (Answers may be different)
test_that("Less than function works with decimals", {
  expect_lt(3.14, 5)
})
```

### Question 5
Add two more test cases to test we already created. This time, add one test that passes and another that will fail.
```{r testthat-q5, eval=FALSE, exercise=TRUE}
## Enter code below:
```

```{r testthat-q5-solution}
## Enter code below: (Answers may be different)
test_that("Less than function works with decimals", {
  expect_lt(3.14, 5)
  expect_lt(3.14, 3.141)
  expect_lt(3.14, 3)
})
```

### Question 6

Finally, add the skip function to skip the test case that fails.
```{r testthat-q6, eval=FALSE, exercise=TRUE}
## Enter code below:
```

```{r testthat-q6-solution}
## Enter code below:
test_that("Less than function works with decimals", {
  expect_lt(3.14, 5)
  expect_lt(3.14, 3.141)
  skip("next test will fail")
  expect_lt(3.14, 3)
})
```

The next three questions will use the code chunk below. We will use the `expect_length` function along with three vectors. 

```{r testthat-q7-9-chunk, eval=FALSE}
vec1 <- c("one")
vec2 <- c("one", "two", "three")
vec3 <- c("one", "two", "three", "four", "six")
test_that("", {
  expect_length(vec1, 1) # 1
  # 2
  expect_length(vec2, 3) # 3
  # 4
  expect_length(vec3, 6) # 5
  # 6
})
```

### Question 7

```{r testthat-q7, echo=FALSE}
question("Which of the lines above will fail its test?",
         answer("1"),
         answer("2"),
         answer("3"),
         answer("4"),
         answer("5", correct = TRUE),
         answer("6"),
         allow_retry = TRUE)
```

### Question 8

```{r testthat-q8, echo=FALSE}
question("Since one of the `expect_length` functions fails its test, which function should we use to not run that test?",
         answer("`skip`", correct = TRUE),
         answer("`avoid`"),
         answer("`next`"),
         answer("None of the above"),
         allow_retry = TRUE)
```

### Question 9

```{r testtthat-q9, echo=FALSE}
question("Which line should we place the skip function to avoid the test case that fails?",
         answer("2"),
         answer("4", correct = TRUE),
         answer("6"),
         allow_retry = TRUE)
```

### Question 10

```{r testthat-q10, echo=FALSE}
order <- c("Run `usethis::use_testthat`", "Edit testing file with different test cases",
           "Create more test files for other functions", 
           "Run your tests using Command/Control + Shift + T")
question_rank("What is the correct order of making a testing set",
              answer(order, correct = T),
              allow_retry = T)
```

## Next Steps

Some next steps and helpful links include:

- The `testthat` website: https://testthat.r-lib.org/
- The testing chapter in the R packages book: https://r-pkgs.org/tests.html

## Questions

1. True or False, `testthat` is located in CRAN?
  a.  True
  b. False
  
2. What folder are tests located in after running `use_testthat`?
  a.  "testthat"
  b. "testing"
  c. "tests"
  d. None of the above
  
3. Which function tests if the output is less than a certain value?
  a.  `expect_lt`
  b. `expect_lte`
  c. `expect_gt`
  d. `expect_gte`
  
4. Which function tests if the output is greater than a certain value?
  a. `expect_lt`
  b. `expect_lte`
  c.  `expect_gt`
  d. `expect_gte`
  
5. Which function can be used to avoid tests?
  a. `avoid`
  b.  `skip`
  c. `null`
  d. `skip_next`
  
6. If I have a function that I think gives an error, which function should I use to ensure an error occurs?
  a. `expect_warning`
  b. `expect_message`
  c.  `expect_error`
  d. None of the above
  
7. True or False, you need to include a description for your tests?
  a.  True
  b. False
  
8. True or False, there is the limit to the amount of tests you can make?
  a. True
  b.  False
  
9. True or False, you must include an error message if you used `expect_error`?
  a. True
  b.  False
  
10. What causes the "argument missing" error in the `test_that` function?
  a.  An argument is missing in the `test_that` function
  b. An argument is missing in the `expect_` function
  c. Too many arguments were used in the `test_that` function
  d. None of the above
  
  
  
  
  
  
